---
title: 'Top Songs Analysis: US, Italy, Japan'
author: "Sergei Korotkikh"
date: "2022-12-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This project is devoted to analyzing the features of Top-50 songs of year 2022 in the United States, Italy and Japan.


## Introduction

The idea for my project was to perform an analysis of the features of top songs of 2022 in different countries. I was interested to see if there are any features that make a song popular, and if such features are the same around the world. I have chosen the three following countries: United States of America, Italy and Japan. When choosing the countries for the analysis, I took into consideration the significance of music market in the country, the region in which the country is located, and the number of "authentic" songs on the playlist of top songs on Spotify. USA and Japan were the two biggest music markets in the world in 2021 [], so I decided to choose them. Other countries (Germany, South Korea, China) either have many American songs or are located in the same region (North America or Asia). So, for my third country I have chosen Italy.

## Data Preparation

For my analysis, I needed to have the data frame for each country's top songs, as well as the features of the songs. To get the data frames, I have used the Spotify API. First, I logged in with the credentials into the system. Then, in the Spotify app, I got the IDs of the playlists (*Top Tracks of 2022 USA*, *Top Songs Italia 2022*, *Top Tracks of 2022 Japan*) and made the request using the API to get the information about the songs in the playlists. Here I encountered the difficulty: the API returned a large list with much unnecessary information about each song. To get the features of each song, I needed only the track ID. Then, using another link of API, I created the data frame with the song features. It was weird for me to see that the values in the table were saved as lists, not as vectors. Luckily, the *unlist* function helped me.

Here is an example of getting the features for top USA songs:

```{r creds, echo=F, warning=F, message=F}
library(tidyverse)
library(httr)
library(stringr)

spotify_cred <- readr::read_csv(file = "credentials/spotify-credentials.csv")
clientID <- spotify_cred$clientID
secret <-  spotify_cred$secret
r <-  POST(
  url = 'https://accounts.spotify.com/api/token',
  accept_json(),
  body = list(grant_type = 'client_credentials',
              client_id = clientID,
              client_secret = secret, 
              content_type = "application/x-www-form-urlencoded"),   
  encode = 'form',
  verbose()
)
cnt <- httr::content(r)
token <- str_c(cnt$token_type, cnt$access_token, sep =' ')

playlist_usa <- 'https://open.spotify.com/playlist/37i9dQZF1DXbJMiQ53rTyJ' %>%
  str_remove(., 'https://open.spotify.com/playlist/')

url_usa <- str_c('https://api.spotify.com/v1/playlists/', playlist_usa, '/tracks')
track_url <- 'https://api.spotify.com/v1/audio-features/'
```

```{r usa example1, warning=F, message=F}
# Loading packages
library(tidyverse)
library(httr)
library(stringr)

# Creating a table with USA song IDs ----

songs_usa <- GET(url = url_usa,  # the link to the playlist
                config = add_headers(authorization = token))

songs_usa <- httr::content(songs_usa)  # extracting the 'large list' to know the way to ID

song_id_usa <- vector(length = 50)  # creating a vector containing songs' IDs

for (i in 1:50) {
  song_id_usa[[i]] = songs_usa$items[[i]]$track$id  # loading the IDs into vector
}

# Creating a table with artist name and song title USA ----

vec_id_usa <- vector(length = 50)  # creating a vector for URL for each song

for (i in 1:50) {
  vec_id_usa[i] <- str_c(track_url, song_id_usa[i])  # creating the URL for each song for API to use
}

names_usa <- data.frame(id = character(),
                    artist = character(),
                    title = character())  # creating a table with song ID, artist and title

for (i in 1:50) {
  names_usa[nrow(names_usa) + 1, ] = c(song_id_usa[i], songs_usa$items[[i]]$track$artists[[1]]$name, songs_usa$items[[i]]$track$name)  # adding the values to the table
}

# Creating a table with songs features USA ----

features_usa <- vector(mode = 'list', length = 50)

for (i in 1:50) {
features_usa[[i]] = GET(url = vec_id_usa[i],
                        config = add_headers(authorization = token))
features_usa[[i]] = httr::content(features_usa[[i]])  # extracting the songs' features into vector
}

songs_f_usa <- do.call('rbind', features_usa)  # creating a table with songs' features
songs_f_usa <- songs_f_usa[, -c(12:16)]  # deleting the unnecessary features
usa <- cbind(names_usa, songs_f_usa)  # creating the table with ID, artist, title and features

# Changing tables' values from lists to vectors ----
for (i in 4:16) {
  usa[, i] <- unlist(usa[, i], use.names = F)
}

# Adding a song position on the chart ----
usa <- add_column(usa, position = 1:50, .after = 1)
```

``` {r usatable, echo=F, warning=F, message=F}
library(knitr)
head(usa[2:7]) %>% kable()
```

I have done the same with the Italian and Japanese playlists.

## Features analysis

Since it is my first research project that involves statistics, I searched for similar analyses to use one as a guideline. I found a research called *"What makes a song trend? Cluster analysis of musical attributes for Spotify top trending songs"* [] in which Top 100 US Trending Songs of 2017 and 2018 were analyzed.

First, I was interested in the descriptive analysis of the means of features for each country. I grouped the songs by countries and counted the means for their features:

```{r means, warning=F, message=F, echo=F}
data <- rio::import('songs/data.csv')
all_means <- data %>% group_by(country) %>% summarise_if(is.numeric, mean)
all_means$position <- NULL  # we do not need the mean for the position
all_means$duration_ms <- apply(all_means[, 'duration_ms'], 1, function(x) (x/60000))  # turning ms to seconds
names(all_means)[names(all_means) == 'duration_ms'] <- 'duration_mins'  # renaming the column
```

```{r datatable, echo=F}
head(all_means[1:6]) %>% kable()
```

Having analyzed the means of the features, I have been able to make the following conclusions. The most popular songs of 2022 in all three countries score high on danceability (>0.6 on the scale from 0 to 1), energy (>0.63) and loudness (>-7 on the scale from -60 to 0 []). At the same time, the songs have low scores in speechiness (<0.12), acousticness (<0.23), instrumentalness (<0.02) and liveness (<0.22). It means that people from the USA, Japan and Italy like energetic songs that are easy to dance to. Also, the songs are not instrumental, they are recorded in the studio, and almost all the words in the songs are sung (not spoken).

I have also paid attention to the differences in the means in particular features between the countries. For example, the *danceability* feature means are almost equal in the US and Italy (0.67790, 0.67388), but is different in Japan (0.60980). t-test (0.05) also proves it:

```{r tdance, echo=F}
filter(data, country %in% c('IT', 'JPN')) %>% t.test(danceability ~ country, .)
```

It means that Japanese people prefer songs that are slightly less danceable than Italians and Americans.

On the other hand, American people enjoy less energetic songs (0.63474) compared to people from Italy and Japan (0.73332, 0.74724):

```{r energybox, echo=F}
data %>% group_by(country) %>% ggplot(., aes(country, energy)) + geom_boxplot() + coord_flip() + theme_bw()
```

ANOVA analysis has also been performed on other variables to see if there are any statistically significant differences between the means of the three countries. The tests on loudness, speechiness, valence and duration of the songs returned the statistically significant p-values. For example, this is the ANOVA test that pays attention to the songs' valence — it shows how positive/cheerful the songs sounds (0 being most sad, 1 being most happy):

```{r valenceaov, echo=F}
summary(aov(valence ~ country, data))
```

Although some tests returned statictically significant values, not all of them are practically significant. For instance, for ANOVA on speechiness p-level is less than 0.01. However, the speechiness values for Italy, Japan and the USA do not vary much: 0.09, 0.07 and 0.11 (on the scale from 0 to 1). Practical significance, as well as statistical one, has been found in valence and duration variables. Japanese people prefer happier songs (0.604) than Italians and Americans (0.509, 0.492). In addition, top songs in Japan are longer (~ 4 minutes) compared to the Italian and American ones (~ 3 and 3.5 minutes).